An assessment of photometric redshift PDF performance in the context of LSST

### 2018-10-12 paper/main.tex
address more referee issues in intro

  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
  However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
  We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
  This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
  We observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
  A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
  % Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
  % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.
  % We present a unified comparison of twelve such codes using simulated LSST photometry, the first in a series of papers on photo-$z$ PDF performance for the LSST Dark Energy Science Collaboration (DESC).
  % In this initial study, we isolate the effects of the estimation procedure by running each code using perfectly representative training data (for machine learning methods) or the true SED template library (for template-fitting methods) and confirm the mutual inconsistency of the resulting photo-$z$ PDF estimates.
  % We apply a comprehensive suite of metrics to an the population of photo-$z$ PDFs as well as point estimates derived thereof and an accepted estimator of the redshift distribution $N(z)$.
  % % The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots, and probability integral transform.
  % We interpret the performance of the codes under each metric in terms of the assumptions underlying the overall approach of the methods and audit the sentivity of the metrics to the failure modes of different techniques, culminating in a discussion predicting the impact of introducing realistic systematics and the direction of attention to the choice of metric when identifying the most appropriate code for a given science use case.

>>>>

  Photometric redshift (photo-$z$) probability distribution functions (PDFs) are a key data product of nearly all upcoming galaxy imaging surveys.
  However, the photo-$z$ PDFs resulting from different techniques are not in general consistent with one another, and an optimal method for obtaining an accurate PDF remains unclear.
  We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}), the first in a planned series of papers testing multiple photo-$z$ codes on simulations of upcoming LSST galaxy photometry catalogues.
  This initial test evaluates photo-$z$ algorithms in the presence of representative training data and in the absence of several common sources of systematic errors that affect the procedures by which photo-$z$ PDFs are derived.
%  The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots and probability integral transform.
  The photo-$z$ PDFs are evaluated using multiple metrics and we observe several trends, including an overall over/under-prediction in the broadness of the PDFs for several of the codes.
  A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.
  % Photometric redshift (photo-$z$) probability distribution functions (PDFs) more completely capture the complex nature of photo-$z$ uncertainties and are thus a key data product of nearly all upcoming galaxy imaging surveys, including the Large Synoptic Survey Telescope (LSST).
  % Photo-$z$ PDFs produced by different algorithms do not in general agree with one another, and no one technique has yet been identified as optimal, leading to a proliferation of photo-$z$ PDF estimation codes.
  % We present a unified comparison of twelve such codes using simulated LSST photometry, the first in a series of papers on photo-$z$ PDF performance for the LSST Dark Energy Science Collaboration (DESC).
  % In this initial study, we isolate the effects of the estimation procedure by running each code using perfectly representative training data (for machine learning methods) or the true SED template library (for template-fitting methods) and confirm the mutual inconsistency of the resulting photo-$z$ PDF estimates.
  % We apply a comprehensive suite of metrics to an the population of photo-$z$ PDFs as well as point estimates derived thereof and an accepted estimator of the redshift distribution $N(z)$.
  % % The photo-$z$ PDFs are evaluated using multiple metrics including the Kolmogorov-Smirnoff statistic, Cramer-von Mises statistic, Anderson-Darling statistic, Kullback-Leibler divergence, $N(z)$ moments, quantile-quantile plots, and probability integral transform.
  % We interpret the performance of the codes under each metric in terms of the assumptions underlying the overall approach of the methods and audit the sentivity of the metrics to the failure modes of different techniques, culminating in a discussion predicting the impact of introducing realistic systematics and the direction of attention to the choice of metric when identifying the most appropriate code for a given science use case.

### 2018-10-12 paper/tex/introduction.tex
address more referee issues in intro

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements on tomographic bin populations.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases seem to need only ensemble redshift information; for instance current weak lensing techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.  However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.  These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.  Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
%In the case of the multiple types of probes of cosmology enabled by the
%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending, and other complications that realistic samples will encounter in estimating photometric redshifts.  This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.

>>>>

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though the initial LSST-DESC requirements are defined for tomographic bin populations.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

For cosmological measurements, certain science cases require redshift information on individual objects, e.~g.~identification of host galaxy redshift for supernova classification, or identifying potential cluster membership.
Other science cases seem to need only ensemble redshift information; for instance many current cosmic shear techniques require only the overall redshift distribution $N(z)$ for tomographic redshift samples.  However, recent data-driven techniques employing hierarchical Bayesian or Gaussian Process methods have emerged that calibrate redshift distributions using individual $p(z)$ estimates \citep[e.~g.~][]{Sanchez:2018}.  These methods assume that the $p(z)$ for each galaxy is an accurate PDF, and such methods break down if this assumption is invalid.  Thus, even methods that seem to need only ensemble $N(z)$ may actually require accurate $p(z)$ in order to meet stringent survey requirements.
%In the case of the multiple types of probes of cosmology enabled by the
%LSST cosmology sample of several billion galaxies, the number of redshift bins and their photo-$z$ requirements vary with the specific probe;  2-point angular correlations benefit from many bins, while weak lensing probes do not (due to the wide lensing kernel).
Large photometric surveys such as LSST must develop algorithms that simultaneously meet the needs of all science cases.
In order to meet these ambitious goals for photo-$z$ accuracy, every aspect of photo-$z$ estimation will have to be optimized: the algorithms employed, both template and machine-learning based (both in design and implementation); the spectroscopic data used as a training set for machine learning algorithms or to estimate template sets and train Bayesian priors; and probabilistic catalogue compression schemes that balance information retention against limited storage resources.

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates \citep{Sanchez:14} and a summary statistic of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.  
This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented in this work are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of representative training sets.
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process of accounting for incomplete training samples, object blending,  the robustness of estimated redshift distributions when combining redshift estimates in tomographic bins, and other complications that realistic samples will encounter in estimating photometric redshifts.  This can be thought of as an initial test under near perfect conditions before further complexities are added in future papers.
